figure sandbox

### 2025-05-16 untargeted/old/paruntargeted_control.Rmd
Moving new untargeted analysis over from ILC

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, purl=FALSE)
options(pillar.sigfig=7)
options(readr.num_columns = 0)
options(readr.show_col_types = FALSE)
library(tidyverse)
library(data.table)
library(arrow)
library(xcms)
library(RaMS)

polarity <- "pos"
# polarity <- "neg"
# amendment <- "Amm"
# amendment <- "Nit"
amendment <- "GMP"
# amendment <- "Arg"
IS_type <- "noIS"
# IS_type <- "wIS"

@@

xcms_file_data <- parametadata %>%
  mutate(samp_group=paste(amendment, depth, startime, timepoint, sep="-"))

mz_group <- function(mz_vals, ppm){
  group_vec <- numeric(length(mz_vals))
  group_num <- 1L
  init_vec <- mz_vals
  names(init_vec) <- seq_along(init_vec)
  while(length(init_vec)>0){
    mz_i <- init_vec[1]
    err <- mz_i*ppm/1000000
    mz_idxs <- init_vec>mz_i-err & init_vec<mz_i+err
    group_vec[as.numeric(names(mz_idxs)[mz_idxs])] <- group_num
    init_vec <- init_vec[!mz_idxs]
    group_num <- group_num+1L
  }
  group_vec
}
trapz <- function(x, y, na.rm=TRUE) {
  y <- y[order(x)]
  x <- x[order(x)]
  m <- length(x)
  xp <- c(x, x[m:1])
  yp <- c(numeric(m), y[m:1])
  n <- 2*m
  p1 <- sum(xp[1:(n-1)]*yp[2:n], na.rm = na.rm) + xp[n]*yp[1]
  p2 <- sum(xp[2:n]*yp[1:(n-1)], na.rm = na.rm) + xp[1]*yp[n]

  return(0.5*(p1-p2))
}
```

@@

```{r making peak_data and best_feat_data}
msnexp_filled <- readRDS(paste0(output_folder, "msnexp_filled.rds"))
peak_data_long <- msnexp_filled %>%
  chromPeaks() %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  mutate(peakidx=row_number())
peak_data <- msnexp_filled %>%
  featureDefinitions() %>%
  as.data.frame() %>%
  select(mzmed, rtmed, npeaks, peakidx) %>%
  rownames_to_column("id") %>%
  unnest_longer(peakidx) %>%
  rename_with(~paste0("feat_", .x), .cols = -peakidx) %>%
  dplyr::rename(feature="feat_id") %>%
  left_join(peak_data_long, by = join_by(peakidx)) %>%
  mutate(filename=basename(fileNames(msnexp_filled))[sample]) %>%
  select(feature, feat_mzmed, feat_rtmed, feat_npeaks, filename,
         mz, mzmin, mzmax, rt, rtmin, rtmax, into, sn, beta_cor, beta_snr) %>%
  arrange(desc(into)) %>%
  group_by(feature, filename) %>%
  slice(1) %>%
  ungroup()
output_csv_write(peak_data)

all_feat_data <- peak_data %>%
  left_join(parametadata) %>%
  filter(samp_type=="Smp") %>%
  group_by(feature) %>%
  summarize(mzmed=unique(feat_mzmed), rtmed=unique(feat_rtmed)/60, n=n(),
            med_cor=median(beta_cor, na.rm=TRUE), 
            med_snr=median(beta_snr, na.rm=TRUE))
output_csv_write(all_feat_data)
# all_feat_data %>%
#   # filter(med_cor>0.75) %>%
#   mutate(n_group=cut(n, breaks = c(0, 2, 10, 40, 60, 62, 100, 130, Inf))) %>%
#   ggplot(aes(label=feature)) +
#   geom_point(aes(x=med_cor, y=med_snr, color=n_group))

@@

best_feat_data <- all_feat_data %>%
  filter(med_cor>0.75 | feature%in%rescue_feats) %>%
  select(-med_cor, -med_snr)
output_csv_write(best_feat_data)
```

@@

```{r plots from mass check and others, eval=FALSE}
best_feat_data %>%
  ggplot() +
  geom_point(aes(x=rtmed, y=mzmed, color=log10(n))) +
  scale_color_viridis_c()

anno_manu %>%
  select(compound_name, feature, mz) %>%
  filter(!feature=="NA") %>%
  left_join(all_feat_data) %>%
  ggplot(aes(label=feature)) +
  geom_point(aes(x=mz, y=mzmed-mz)) +
  geom_abline(slope = 1, intercept = 0)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, purl=FALSE)
options(pillar.sigfig=7)
options(readr.num_columns = 0)
options(readr.show_col_types = FALSE)
library(tidyverse)
library(data.table)
library(arrow)
library(xcms)
library(MSnbase)
library(RaMS)
library(squallms)

polarity <- "pos"
# polarity <- "neg"
amendment <- "Amm"
# amendment <- "Nit"
# amendment <- "GMP"
# amendment <- "Arg"
IS_type <- "noIS"
# IS_type <- "wIS"

@@

xcms_file_data <- parametadata %>%
  mutate(samp_group=paste(amendment, depth, startime, timepoint, sep="-"))
```

@@

```{r making peak_data and best_feat_data}
msnexp_filled <- readRDS(paste0(output_folder, "msnexp_filled.rds"))

peak_data <- makeXcmsObjFlat(msnexp_filled, revert_rts = FALSE) %>%
  group_by(filename, feature) %>%
  arrange(desc(into)) %>%
  slice(1) %>%
  ungroup()
output_csv_write(peak_data)

all_feat_data <- peak_data %>%
  select(-filepath) %>%
  left_join(parametadata) %>%
  filter(samp_type=="Smp") %>%
  group_by(feature) %>%
  summarize(mzmed=unique(feat_mzmed), rtmed=unique(feat_rtmed)/60, n=n(),
            med_cor=median(beta_cor, na.rm=TRUE), 
            med_snr=median(beta_snr, na.rm=TRUE))
output_csv_write(all_feat_data)

@@

```

```{r}
# best_feat_data <- all_feat_data %>%
#   filter(med_cor>0.75 | feature%in%rescue_feats) %>%
#   select(-med_cor, -med_snr)
# output_csv_write(best_feat_data)

library(squallms)
pk_data <- makeXcmsObjFlat(msnexp_filled, revert_rts = FALSE) %>%
  filter(str_detect(filename, "Poo"))
msdata <- grabMSdata(unique(pk_data$filepath))
feat_metrics <- extractChromMetrics(pk_data, recalc_betas = TRUE, verbosity = 2, 
                                    ms1_data = msdata$MS1)
class_labels <- labelFeatsLasso(pk_data, verbosity=1, ms1_data = msdata$MS1)
good_cmpds <- logModelFeatQuality(feat_metrics, class_labels, likelihood_threshold=0.8)
cmpds_keep <- c(good_cmpds, rescue_feats)

best_feat_data <- all_feat_data %>%
  filter(feature%in%cmpds_keep)
output_csv_write(best_feat_data)
```

```{r}
manual_labels <- labelFeatsManual(peak_data, existing_labels = class_labels, 
                                  ms1_data = msdata$MS1, selection = "Labeled")
table(manual_labels, class_labels)

```

@@

```{r plots from mass check and others, eval=FALSE}
best_feat_data %>%
  ggplot() +
  geom_point(aes(x=rtmed, y=mzmed, color=n)) +
  scale_color_viridis_c()

anno_manu %>%
  select(compound_name, feature, mz) %>%
  filter(!feature=="NA") %>%
  left_join(all_feat_data) %>%
  ggplot(aes(label=feature)) +
  geom_point(aes(x=mz, y=mzmed)) +
  geom_abline(slope = 1, intercept = 0)

### 2025-05-16 untargeted/old/paruntargeted_sandbox.Rmd
Moving new untargeted analysis over from ILC

all_areas <- map(c("pos", "neg"), function(pol_i){
  pol_i_areas <- map(c("Amm", "Nit", "GMP", "Arg"), function(amendment_i){
    file_path <- paste0("G:/My Drive/PARAGON/untargeted/", amendment_i, "_", 
                        pol_i, "_noIS_output/iso_areas.csv")
      read_csv(file_path) %>%
      mutate(amendment=amendment_i)
  }) %>% bind_rows() %>%
    mutate(polarity=pol_i)
}) %>% bind_rows()
all_rfs <- all_areas %>% 
  group_by(amendment, polarity) %>%
  filter(samp_type=="Std") %>%
  filter(labeled_n_count==0) %>%
  left_join(all_stans %>% select(compound_name, mix, conc_um, polarity)) %>%
  mutate(samp_type=str_extract(filename, paste0(mix, "InMatrix|H2OInMatrix"))) %>%
  drop_na() %>%
  mutate(samp_type=str_remove(samp_type, "(?<=Mix)\\d")) %>%
  select(compound_name, samp_type, area, conc_um) %>%
  pivot_wider(names_from=samp_type, values_from = area, values_fn = mean) %>%
  mutate(rf=(MixInMatrix-H2OInMatrix)/conc_um) %>%
  select(compound_name, rf) %>%
  filter(!is.na(rf)) %>%
  filter(rf>0)
all_concs <- all_areas %>%
  right_join(all_rfs) %>%
  mutate(uM_in_vial=area/rf) %>%
  mutate(uM_in_env=(uM_in_vial*(400/1e6))/2) %>%
  mutate(nM_in_env=uM_in_env*1e3) %>%
  select(compound_name, filename, amendment, polarity, labeled_n_count, nM_in_env)
```

@@

>>>>

all_areas <- map(c("pos", "neg"), function(pol_i){
  pol_i_areas <- map(c("Amm", "Nit", "GMP", "Arg"), function(amendment_i){
    file_path <- paste0("C:/Users/willi/Documents/PARAGON/untargeted/", amendment_i, "_", 
                        pol_i, "_noIS_output/iso_areas.csv")
      read_csv(file_path) %>%
      mutate(amendment=amendment_i)
  }) %>% bind_rows() %>%
    mutate(polarity=pol_i)
}) %>% bind_rows()
all_rfs <- all_areas %>% 
  group_by(amendment, polarity) %>%
  filter(samp_type=="Std") %>%
  filter(labeled_n_count==0) %>%
  left_join(all_stans %>% select(compound_name, mix, conc_um, polarity)) %>%
  mutate(samp_type=str_extract(filename, paste0(mix, "InMatrix|H2OInMatrix"))) %>%
  drop_na() %>%
  mutate(samp_type=str_remove(samp_type, "(?<=Mix)\\d")) %>%
  select(compound_name, samp_type, area, conc_um) %>%
  pivot_wider(names_from=samp_type, values_from = area, values_fn = mean) %>%
  mutate(rf=(MixInMatrix-H2OInMatrix)/conc_um) %>%
  select(compound_name, rf) %>%
  filter(!is.na(rf)) %>%
  filter(rf>0)
all_concs <- all_areas %>%
  right_join(all_rfs) %>%
  mutate(uM_in_vial=area/rf) %>%
  mutate(uM_in_env=(uM_in_vial*(400/1e6))/2) %>%
  mutate(nM_in_env=uM_in_env*1e3) %>%
  select(compound_name, filename, amendment, polarity, labeled_n_count, nM_in_env)
```

@@

# Fresh start

```{r}
msnexp_filled <- readRDS("untargeted/Amm_pos_noIS_output/msnexp_filled.rds")
summarized_feats <- msnexp_filled %>%
  makeXcmsObjFlat(revert_rts = FALSE) %>%
  summarise(feat_mzmed=unique(feat_mzmed), feat_rtmed=unique(feat_rtmed),
            feat_into=mean(into, na.rm=TRUE), mean_cor=mean(beta_cor, na.rm=TRUE),
            mean_snr=mean(beta_snr, na.rm=TRUE), n=n(), .by=feature)
summarized_feats %>%
  ggplot() +
  geom_point(aes(x=feat_rtmed, y=feat_mzmed, color=log10(feat_into)))
summarized_feats %>%
  ggplot() +
  geom_point(aes(x=mean_cor, y=mean_snr, color=log10(feat_into)))
summarized_feats %>%
  mutate(plot_col=feat_mzmed%between%pmppm(118.0865, 10)) %>%
  ggplot() +
  geom_point(aes(x=mean_cor, y=mean_snr, color=plot_col))
```

```{r}
library(squallms)
ms_peak_data <- msnexp_filled %>%
  makeXcmsObjFlat(revert_rts = FALSE) %>%
  filter(str_detect(filename, "Poo")) %>%
  mutate(filepath=str_replace(filepath, "G:\\\\My Drive\\\\", "C:\\\\Users\\\\willi\\\\Documents\\\\"))
labeled_feats <- squallms::labelFeatsLasso(ms_peak_data)

good_feats <- msnexp_filled %>%
  makeXcmsObjFlat(revert_rts = FALSE) %>%
  summarise(med_cor=median(beta_cor, na.rm = TRUE), med_snr=median(beta_snr, na.rm = TRUE), .by=feature) %>%
  squallms::logModelFeatQuality(feature_labels = labeled_feats, likelihood_threshold = 0.5)
```

```{r}
msnexp_filled %>%
  makeXcmsObjFlat(revert_rts = FALSE) %>%
  summarise(feat_mzmed=unique(feat_mzmed), feat_rtmed=unique(feat_rtmed),
            feat_into=mean(into, na.rm=TRUE), mean_cor=mean(beta_cor, na.rm=TRUE),
            mean_snr=mean(beta_snr, na.rm=TRUE), n=n(), .by=feature) %>%
  filter(feature%in%good_feats) %>%
  ggplot() +
  # geom_point(aes(x=feat_rtmed, y=feat_mzmed, color=log10(feat_into)))
  geom_point(aes(x=mean_cor, y=mean_snr, color=log10(feat_into)))
```
